Covering Basis: Comprehensive Model Agnostic Probes of the Binary Black Hole Population

### 2022-10-04 src/tex/conclusion.tex
more comments fixing up text/wording/grammar

% mass dist
The results presented in section \ref{sec:mass_dist} illustrate a much wider mass distribution than inferred with power-law based models in \citet{o3b_astro_dist}. 
While isolated formation is able to predict the $10\msun$ peak \bruce{cite}, cluster and dynamical formation scenarios struggle to predict a peak in the BH mass distribution less than 
$15-20\msun$ \bruce{cite}. Globular cluster formation is expected to produce more top-heavy mass distributions than isolated and recent studies have shown suppressed BBH merger rates 
at lower ($m\leq15\msun$) masses when compared to predictions from the isolated channel \bruce{cite}. BBHs that form near active galactic nuclei (AGN) can preferentially produce higher 
mass black holes \bruce{cite}. We do not find any evidence for a truncation or rapid decline in the merger rate as a function of mass, that stellar evolution theory predicts 
due to pair-instability supernovae (PISNe) \citep{Heger_2002,PISN_Woosley,Heger_2003,Spera_2017}. The original motivation for the peak in the \textsc{PowerlawPeak} model \citep{Talbot_2018} was to represent a possible ``pileup'' of 
masses just before such the truncation, since massive stars just light enough to avoid PISN will shed large amounts of mass in a series of ``pulses'' before collapsing to BHs in a process called 
pulsational pair-instability supernova (PPISN) \citep{Woosley_2017,Woosley_2019,Farmer_2019}. While the predictions of the mass scale where pair-instability kicks in are uncertain and depend on not well understood physics, 
like nuclear reaction rates of carbon and oxygen in the core of stars, models have a hard time predicting this peak to be lower than $m\sim40\msun$ \citep{Belczynski_2016,Marchant_2019,Renzo_2020,Farmer_2019,Farmer_2020}. The lack of a truncation could 
point towards a higher prevalence of dynamical processes that can produce black holes in mass ranges stellar collapse cannot, such as hierarchical mergers of BHs \citep{Fishbach_2017,Doctor_2020,Kimball_genealogy,kimball2020evidence,doctor2021black,Fishbach_2022}, 
very low metallicity population III stars \citep{Belczynski_2020,Farrell_2020}, new beyond the standard model physics\citep{Croon_newphysics,Sakstein_2020}, or black hole accretion of BHs in gaseous environments such as AGNs \citep{Secunda_2020,McKernan_2020,cruzosorio2021gw190521}.

% spin dist
Section \ref{sec:spin_dist} focused on inferences of the spin distributions of black holes, observing evidence of spin misalignment, spin anti-alignment, and suppressed support 
for exactly aligned systems. These point towards a significant contribution of the population being formed through dynamical formation processes, agreeing with 
conclusions drawn about the mass distribution inference of section \ref{sec:mass_dist}. While field formation is expected to produce systems with preferentially 
aligned spins due to tidal interactions, but observational evidence suggests that tides may not be able to re-align spins in all systems as some 
isolated population models assume. Additionally, because of uncertain knowledge of supernovae kicks, isolated formation can produce systems with negative but small effective spins. 
We report an effective spin distribution that is not symmetric about zero, disfavoring a scenario in which all BBHs are formed dynamically. Following the rules in \citet{Fishbach_2022}, 
we place conservative upper bounds on the fraction of hierarchical mergers $f_\mathrm{HM}$ and fraction of dynamically formed BBHs, $f_\mathrm{dyn}$ with 
the M-Spline $\chi_\mathrm{eff}$ model constraining $f_\mathrm{HM} < $\result{$\macros[ChiEffective][chieff][frac_hm][10th percentile]$} 
and $f_\mathrm{dyn} < $\result{$\macros[ChiEffective][chieff][frac_dyn][10th percentile]$} at 90\% credibility.

>>>>

% mass dist
The results presented in section \ref{sec:mass_dist} illustrate a much wider mass distribution than inferred with power-law based models in \citet{o3b_astro_dist}. 
While isolated formation is able to predict the $10\msun$ peak \bruce{cite}, cluster and dynamical formation scenarios struggle to predict a peak in the BH mass distribution less than 
$15-20\msun$ \bruce{cite}. Globular cluster formation is expected to produce more top-heavy mass distributions than isolated and recent studies have shown suppressed BBH merger rates 
at lower ($m\leq15\msun$) masses when compared to predictions from the isolated channel \bruce{cite}. BBHs that form near active galactic nuclei (AGN) can preferentially produce higher 
mass black holes \bruce{cite}. We do not find any evidence for a truncation or rapid decline in the merger rate as a function of mass, that stellar evolution theory predicts 
due to pair-instability supernovae (PISNe) \citep{Heger_2002,PISN_Woosley,Heger_2003,Spera_2017}. The original motivation for the peak in the \textsc{PowerlawPeak} model \citep{Talbot_2018} was to represent a possible ``pileup'' of 
masses just before such truncation, since massive stars just light enough to avoid PISN will shed large amounts of mass in a series of ``pulses'' before collapsing to BHs in a process called 
pulsational pair-instability supernova (PPISN) \citep{Woosley_2017,Woosley_2019,Farmer_2019}. While the predictions of the mass scale where pair-instability kicks in are uncertain and depend on poorly understood physics, 
like nuclear reaction rates of carbon and oxygen in the core of stars, models have a hard time predicting this peak to be lower than $m\sim40\msun$ \citep{Belczynski_2016,Marchant_2019,Renzo_2020,Farmer_2019,Farmer_2020}. The lack of a truncation could 
point towards a higher prevalence of dynamical processes that can produce black holes in mass ranges stellar collapse cannot, such as hierarchical mergers of BHs \citep{Fishbach_2017,Doctor_2020,Kimball_genealogy,kimball2020evidence,doctor2021black,Fishbach_2022}, 
very low metallicity population III stars \citep{Belczynski_2020,Farrell_2020}, new beyond the standard model physics\citep{Croon_newphysics,Sakstein_2020}, or black hole accretion of BHs in gaseous environments such as AGNs \citep{Secunda_2020,McKernan_2020,cruzosorio2021gw190521}.

% spin dist
Section \ref{sec:spin_dist} focused on inferences of the spin distributions of black holes, observing evidence of spin misalignment, spin anti-alignment, and suppressed support 
for exactly aligned systems. These point towards a significant contribution of the population being formed through dynamical formation processes, agreeing with 
conclusions drawn about the mass distribution inference of section \ref{sec:mass_dist}. While field formation is expected to produce systems with preferentially 
aligned spins due to tidal interactions, observational evidence suggests that tides may not be able to re-align spins in all systems as some 
isolated population models assume. Additionally, because of uncertain knowledge of supernovae kicks, isolated formation can produce systems with negative but small effective spins. 
We report an effective spin distribution that is not symmetric about zero, disfavoring a scenario in which all BBHs are formed dynamically. Following the rules in \citet{Fishbach_2022}, 
we place conservative upper bounds on the fraction of hierarchical mergers $f_\mathrm{HM}$ and fraction of dynamically formed BBHs, $f_\mathrm{dyn}$ with 
the M-Spline $\chi_\mathrm{eff}$ model constraining $f_\mathrm{HM} < $\result{$\macros[ChiEffective][chieff][frac_hm][10th percentile]$} 
and $f_\mathrm{dyn} < $\result{$\macros[ChiEffective][chieff][frac_dyn][10th percentile]$} at 90\% credibility.

### 2022-10-04 src/tex/intro.tex
more comments fixing up text/wording/grammar

Observations of gravitational waves (GWs) from compact binary coalescences (CBCs) are becoming a regular occurrence, 
recently producing catalogs of 90 such detections \citep{GWTC1,gwtc2,GWTC3}. With this growing catalog size, we are steadily approaching a more complete  
understanding of the population properties of CBCs, just 7 years after the first observation \citep{o1o2_pop,o3a_pop,o3b_astro_dist}. 
In the coming fourth observing run of the LIGO-Virgo-KAGRA (LVK) collaboration detectors \citep{aLIGO,aVIRGO}, with the increased sensitivities, 
there could be a detection as often as every day \citep{LVK_prospects}. This large influx of incoming data will drastically increase what we can 
learn by combining all observations in order to study the compact object population. The history of these dense objects is imprinted on the properties 
of the population  of CBCs, allowing us to learn about their formation and evolution by the last few seconds of gravitational radiation emitted \citep{Zevin_2017}. 
The population distributions of merging compact objects inferred from gravitational-wave observations have been shown to be particularly useful in order to
measure cosmological parameters \citep{Farr_2019HUB,gwtc3_cosmo}, constrain modified gravitational wave propagation \citep{ModGWProp,ModGWProp2}, 
constrain a running Planck mass \citep{Lagos_runningPlanckMass}, look for evidence of ultralight bosons through superradiance \citep{Ng_Boson2021,GWTC2_superradiance_Ng}, 
constrain stellar nuclear reaction rates \citep{Farmer_2019,Farmer_2020}, look for primordial black holes \citep{Ng_2021,KenNgPBH2022}, 
and to constrain physics of neutron stars \citep{Golomb_EOS,LandryRead_NS_Masses2021}. Through a better understanding of the mass, spin, and redshift distributions of 
compact binaries that will come with the increased catalog size, one can probe a wide range of different physical phenomena with even greater fidelity.

Studies using flexible methods to analyze the binary black hole (BBH) mass distribution with GWTC-2 found signs of structure 
beyond the first reported feature found at $m_1\sim35\,\msun$ \citep{Talbot_2018,o3a_pop,Tiwari_2021_b,Edelman_2022ApJ}. The evidence for this structure, 
such as the peak at $m_1\sim10\,\msun$, has only increased after analyzing the following catalog, GWTC-3 \citep{o3b_astro_dist,Tiwari_2022ApJ}. 
This substructure was first confidently found with multiple different flexible or non-parametric models that are data-driven with little 
bias on the functional form. Conversely, the other canonical approach is to use simple parametric models that aim to describe the data
in the simplest way, building in astrophysically motivated priors, followed by sequentially adding complexities only as the data demands it. While in data-scarce, prior dominated
environments, this simple approach is more powerful. However, as we go forward into the data-rich catalog era, flexible models will be necessary to discover new features that may be in
the data, which quickly becomes unwieldy with the sequential approach. Non-parametric or semi-parametric methods are particularly useful in looking for unexpected features in the data, finding where 
our simpler parametric models may be failing to capture meaningful features in the data. Development and robust testing of such methods in the short term will provide
knock on effects as the size of our catalogs increase, and GW astronomy enters its data-dominated era.

The spin distributions of BBHs are particularly interesting since the two main formation scenarios produce distinctly different spin distributions \bruce{cite}. 
Isolated or field formation scenarios predict component spins that are preferentially aligned with the binaries orbital angular momentum, although some small 
misalignment can occur depending on the nature of the supernova kicks as each star collapses to a compact object \bruce{cite}. Dynamical formation scenarios may 
occur in dense environments where many-body interactions between a large number compact objects occur during the binary formation hardening the binary so that it 
will merge within a Hubble time. Thus, the dynamical channel produces an isotropic spin tilt distribution \bruce{cite}. BBH spins have also been of controversial 
interest recently, with different parametric approaches to modeling the spin distribution coming to different conclusions. The studies have disagreed on the possible existence 
of a significant zero-spinning subpopulation, along with the presence of significant spin misalignment (i.e. $\cos{\theta_i} < 0.0$) \citep{o3b_astro_dist,RouletGWTC2Pop,BuildBetterSpinModels,GWTC3MonashSpin,Callister_NoEvidence}. 
Another study recently showed that the spin tilt distribution inference is sensitive to modeling choices and may not peak at perfectly aligned spins, as is often assumed \citep{spinitasyoulike}. 
Other approaches to building improved models for BBH spins again, as with the mass models, focus on iterating through different simple parametric descriptions with 
increasing complexities \citep{BuildBetterSpinModels,Callister_NoEvidence,spinitasyoulike}. The spin distributions offer a nice use case for flexible data-driven models, as 
we present in this manuscript, both to help alleviate the current disputes in the literature and because there has been little work on flexible model development for the 
spin distribution of BBHs beyond the aligned spin dimension thus far \citep{Tiwari_2021_a,Tiwari_2021_b}.

Polynomial splines have been applied to success across different areas of gravitational-wave astronomy. They have been used to model the gravitational-wave data noise spectrum, 
detector calibration uncertainties, coherent gravitational waveform deviations, and modulations to a power law mass distribution \citep{Littenberg_2015,Edwards_2018,B_Farr_etal_2014,Edelman_2021,Edelman_2022ApJ}
In this paper we highlight how the use of basis-splines can provide a powerful non-parametric modeling approach to the astrophysical distributions of compact 
binaries. We illustrate how one can efficiently model both the mass and spin distributions of merging compact binaries in GWTC-3 with basis splines to infer CBC population properties using 
hierarchical Bayesian inference. We discuss our results in the context of current literature on compact object populations and how this method complements the simpler lower 
dimensional parametric models in the short run, and will become necessary with future catalogs. This data-driven approach will provide checks of our understanding by finding more 
subtle features, that may not be expected, should they appear in with more observations. The rest of this manuscript is structured by first going over the background of basis splines in 
section \ref{sec:basis_splines}, followed by presenting the results of our extensive data-driven study of the mass and spin distributions of BBHs in GWTC-3 section \ref{sec:results}. We 
discuss our results and their astrophysical implications in section \ref{sec:astrodiscussion}, finishing with our conclusions in section \ref{sec:conclusion}.

>>>>

Observations of gravitational waves (GWs) from compact binary coalescences (CBCs) are becoming a regular occurrence, 
recently producing catalogs of 90 such detections \citep{GWTC1,gwtc2,GWTC3}. With this growing catalog size, we are steadily approaching a more complete  
understanding of the population properties of CBCs, just 7 years after the first observation \citep{o1o2_pop,o3a_pop,o3b_astro_dist}. 
Due to their increased sensitivity, detections could occur as often as every day in the upcoming fourth observing run of the LIGO-Virgo-KAGRA (LVK) collaboration 
detectors \citep{aLIGO,aVIRGO,LVK_prospects}. This large influx of data will drastically improve our ability to study the population of observed CBCs by 
combining the insights from the entire catalog. The history of these dense objects is imprinted on the properties 
of the population  of CBCs, allowing us to learn about their formation and evolution from the last few seconds of gravitational radiation emitted \citep{Zevin_2017}. 
The population distributions of merging compact objects inferred from gravitational-wave observations have been shown to be particularly useful in order to
measure cosmological parameters \citep{Farr_2019HUB,gwtc3_cosmo}, constrain modified gravitational wave propagation \citep{ModGWProp,ModGWProp2}, 
constrain a running Planck mass \citep{Lagos_runningPlanckMass}, look for evidence of ultralight bosons through superradiance \citep{Ng_Boson2021,GWTC2_superradiance_Ng}, 
constrain stellar nuclear reaction rates \citep{Farmer_2019,Farmer_2020}, look for primordial black holes \citep{Ng_2021,KenNgPBH2022}, 
and to constrain physics of neutron stars \citep{Golomb_EOS,LandryRead_NS_Masses2021}. Through a better understanding of the mass, spin, and redshift distributions of 
compact binaries that will come with the increased catalog size, one can probe a wide range of different physical phenomena with even greater fidelity.

Studies using flexible methods to analyze the binary black hole (BBH) mass distribution with GWTC-2 found signs of structure beyond the first reported feature found at 
$m_1\sim35\,\msun$ \citep{Talbot_2018,o3a_pop,Tiwari_2021_b,Edelman_2022ApJ}. The evidence for this structure, 
such as the peak at $m_1\sim10\,\msun$, only grew after analyzing the latest catalog, GWTC-3, with the same models \citep{o3b_astro_dist,Tiwari_2022ApJ}, illustrating  
the usefulness of data-driven methods even with the relatively small current catalog size. Canonically, the approach to constructing population models has been 
to use simple parametric models that aim to describe the data in the simplest way, build in astrophysically motivated priors, then sequentially add in 
complexities only as the data demands. This simple approach is powerful in data-scarce, prior dominated environments; however, as we go forward into 
the data-rich catalog era, flexible methods such as non-parametric modeling will be necessary to discover new features in the data that can quickly 
become unwieldy to incorporate into the parametric approach. As opposed to parametric models, flexible and non-parametric models are data-driven and 
contribute little bias to functional form. Non-parametric or semi-parametric methods are particularly useful to search for unexpected features in the 
data, such as meaningful features that parametric models may fail to capture. Development and robust testing of such flexible methods in the short 
term will only provide knock on effects when the time comes to analyze the next catalog of CBC detections.

The spin distributions of BBHs are particularly interesting since the two main formation scenarios produce distinctly different spin distributions \bruce{cite}. 
Isolated or field formation scenarios predict component spins that are preferentially aligned with the binaries orbital angular momentum, although some small 
misalignment can occur depending on the nature of the supernova kicks as each star collapses to a compact object \bruce{cite}. Dynamical formation scenarios may occur in 
dense environments where many-body interactions between many compact objects cause a binary to harden so that it will merge within a Hubble time. Thus, the dynamical 
channel produces an isotropic spin tilt distribution \bruce{cite}. BBH spins have also been of controversial interest recently, with different parametric approaches to modeling 
the spin distribution coming to different conclusions. The studies have disagreed on the possible existence of a significant zero-spinning subpopulation, along with the presence of 
significant spin misalignment (i.e. $\cos{\theta_i} < 0.0$) \citep{o3b_astro_dist,RouletGWTC2Pop,BuildBetterSpinModels,GWTC3MonashSpin,Callister_NoEvidence}. 
Another study recently showed that the spin tilt distribution inference is sensitive to modeling choices and may not peak at perfectly aligned spins, as is often assumed \citep{spinitasyoulike}. 
Other approaches to building improved models for BBH spins again, as with the mass models, focus on iterating through different simple parametric descriptions with 
increasing complexities \citep{BuildBetterSpinModels,Callister_NoEvidence,spinitasyoulike}. The spin distributions offer a nice use case for flexible data-driven models, as 
we present in this manuscript, both to help alleviate the current disputes in the literature and because there has been little work on flexible model development for the 
spin distribution of BBHs beyond the aligned spin dimension thus far \citep{Tiwari_2021_a,Tiwari_2021_b}.

Polynomial splines have been applied to success across different areas of gravitational-wave astronomy. They have been used to model the gravitational-wave data noise spectrum, 
detector calibration uncertainties, coherent gravitational waveform deviations, and modulations to a power law mass distribution \citep{Littenberg_2015,Edwards_2018,B_Farr_etal_2014,Edelman_2021,Edelman_2022ApJ}
In this paper we highlight how the use of basis-splines can provide a powerful non-parametric modeling approach to the astrophysical distributions of compact 
binaries. We illustrate how one can efficiently model both the mass and spin distributions of merging compact binaries in GWTC-3 with basis splines to infer CBC population properties using 
hierarchical Bayesian inference. We discuss our results in the context of current literature on compact object populations and how this method complements the simpler lower 
dimensional parametric models in the short run, and will become necessary with future catalogs. Should they appear with more observations, this data-driven approach will provide checks of 
our understanding by uncovering more subtle -- potentially unexpected -- features. The rest of this manuscript is structured as follows: a description of the background of 
basis splines in section \ref{sec:basis_splines}, followed by a presentation of the results of our extensive, data-driven study of the mass and spin distributions of BBHs in GWTC-3 in section 
\ref{sec:results}. We then discuss these results and their astrophysical implications in section \ref{sec:astrodiscussion} and finish with our conclusions in section \ref{sec:conclusion}.

### 2022-10-04 src/tex/methods.tex
more comments fixing up text/wording/grammar

\noindent This is known as the ``B-Spline'' basis after it's inventor de Boor \citep{deBoor78}. The power of basis splines
comes from the fact that one only has to do the somewhat-expensive interpolation once for each set of points at which the spline is evaluated. 
This provides a considerable computational speedup since each evaluation of the spline function becomes a simpler operation that is just
a dot product of a matrix and a vector. This straightforward operation is also ideal for optimizations from the use GPU accelerators, 
enabling expensive Markov chain Monte Carlo (MCMC) based analyses with hundreds of parameters \bruce{citation?}. 
Basis splines can easily be generalized to their two-dimensional analog, producing tensor product basis splines that, 
with this computational advantage, allow for high fidelity modeling of two-dimensional spline functions.

>>>>

\noindent This is known as the ``B-Spline'' basis after it's inventor de Boor \citep{deBoor78}. The power of basis splines
comes from the fact that one only has to do the somewhat-expensive interpolation once for each set of points at which the spline is evaluated. 
This provides a considerable computational speedup as each evaluation of the spline function becomes a simpler operation: a dot product of a 
matrix and a vector. This straightforward operation is also ideal for optimizations from the use of GPU accelerators, 
enabling expensive Markov chain Monte Carlo (MCMC) based analyses with hundreds of parameters \bruce{citation?}. 
Basis splines can easily be generalized to their two-dimensional analog, producing tensor product basis splines that, 
with this computational advantage, allow for high fidelity modeling of two-dimensional spline functions.

### 2022-10-04 src/tex/ms.tex
Possible slight title change? highlighting first comprehensive data driven study
new title?
more comments fixing up text/wording/grammar

\usepackage{showyourwork}
%\usepackage[perpage]{footmisc}
%\usepackage{colortbl}
%\usepackage{lineno}
%\linenumbers
\usepackage{bm}
\usepackage{amsmath}
\usepackage{appendix}
%\usepackage[margin=10pt]{caption}
\input{macros}

%\definecolor{Gray}{gray}{0.9}

@@

\title{Covering Your Basis: Model Agnostic Probes of the Binary Black Hole Population}

\author{Bruce Edelman}
\email{bedelman@uoregon.edu}
\affiliation{Institute  for  Fundamental  Science, Department of Physics, University of Oregon, Eugene, OR 97403, USA}

\author{Ben Farr}
\affiliation{Institute  for  Fundamental  Science, Department of Physics, University of Oregon, Eugene, OR 97403, USA}

\author{Zoheyr Doctor}
\affiliation{Center for Interdisciplinary Exploration and Research in Astrophysics (CIERA), Department of Physics and Astronomy, Northwestern
University, Evanston, IL 60201, USA}

%\author{Jaxen Godfrey}
%\affiliation{Institute  for  Fundamental  Science, Department of Physics, University of Oregon, Eugene, OR 97403, USA}

%\author{Alexis Vives}
%\affiliation{Institute  for  Fundamental  Science, Department of Physics, University of Oregon, Eugene, OR 97403, USA}

%\author{Jon Hornung}
%\affiliation{Institute  for  Fundamental  Science, Department of Physics, University of Oregon, Eugene, OR 97403, USA}
%\affiliation{Thurston High School, Springfield, OR 9747, USA}

\begin{abstract}                 
We introduce a non-parametric model built from basis splines, and use it to infer the astrophysical distribution of 
binary black hole (BBH) masses and spins using the 69 confident observations in GWTC-3, in the most comprehensive data-driven
investigation of the compact object population. We report the same features previously recovered 
with similarly flexible models focusing solely on the mass distribution, most notably the peaks in merger rates at $\sim10\msun$ and $\sim35\msun$. 
Our study provides the first non-parametric investigation into BBH spin distribution, while simultaneously modeling the mass distribution in the same way, 
showing signs that the spin tilt distributions may not peak at $\cos(\theta_i) = 1$ as previous models assumed. We find broad agreement with the 
accepted conclusions of the spin magnitude distribution: the majority of BBH spins are small ($a<0.5$), the distribution peaks at $a\sim0.2$, 
and there is mild support for a non-spinning subpopulation, which may become resolvable with larger catalogs. 
We also perform a data-driven test of whether the data prefers the spin distributions to be identical for both the more and less massive binary components, 
finding no strong sign of differing distributions. When modeling the $\chi_\mathrm{eff}$ distribution with our model we find similar 
conclusions as previously reported, the distribution is consistent with a non-isotropic distribution (that peaks at $\chi_\mathrm{eff}=\,$\result{$\CIPlusMinus{\macros[ChiEffective][chieff][PeakChiEff]}$}), 
showing considerable support ($f_{\chi_\mathrm{eff}<0}=\,$\result{$\CIPlusMinus{\macros[ChiEffective][chieff][FracBelow0]}$}) for negative effective spins or misaligned systems.
We conclude by discussing how non-parametric methods in gravitational-wave population inference are poised to provide a useful complement to 
the parametric approach going into the data-rich catalog era of gravitational-wave astronomy.
\end{abstract}

@@

\section{Acknowledgements}\label{sec:acknowledments}
We thank Will Farr, Tom Callister, Amanda Farah and Maya Fishbach for useful discussions during the preparation this manuscript, 
along with helpful comments on an early draft by Tom Callister.
This research has made use of data, software and/or web tools obtained from the Gravitational Wave Open Science Center 
(\url{https://www.gw-openscience.org/}), a service of LIGO Laboratory, the LIGO Scientific Collaboration and the Virgo Collaboration. 
The authors are grateful for computational resources provided by the LIGO Laboratory and supported by National Science Foundation Grants PHY-0757058 and PHY-0823459.  
This work benefited from access to the University of Oregon high performance computer, Talapas. This material is based upon work supported 
in part by the National Science Foundation under Grant PHY-1807046 and work supported by NSF's LIGO Laboratory which is a major facility 
fully funded by the National Science Foundation.
\software{
\textsc{Astropy}~\citep{2018AJ....156..123A},
\textsc{NumPy}~\citep{harris2020array},
\textsc{Matplotlib}~\citep{Hunter:2007},
\textsc{Jax}~\citep{jax},
\textsc{NumPyro}~\citep{pyro,numpyro},
\textsc{GWPopulation}~\citep{Talbot_2019},
\textsc{Bilby}~\citep{Ashton_2019,bilby_gwtc1},
\textsc{Dynesty}~\citep{Speagle_2020}
}
\bibliography{bib}{}
\bibliographystyle{aasjournal}

>>>>

\usepackage{showyourwork}
\usepackage{bm}
\usepackage{amsmath}
\usepackage{appendix}
\input{macros}

@@

\title{Covering Basis: Comprehensive Model Agnostic Probes of the Binary Black Hole Population}

\author{Bruce Edelman}
\email{bedelman@uoregon.edu}
\affiliation{Institute  for  Fundamental  Science, Department of Physics, University of Oregon, Eugene, OR 97403, USA}
\author{Ben Farr}
\affiliation{Institute  for  Fundamental  Science, Department of Physics, University of Oregon, Eugene, OR 97403, USA}
\author{Zoheyr Doctor}
\affiliation{Center for Interdisciplinary Exploration and Research in Astrophysics (CIERA), Department of Physics and Astronomy, Northwestern
University, Evanston, IL 60201, USA}
%\author{Jaxen Godfrey}
%\affiliation{Institute  for  Fundamental  Science, Department of Physics, University of Oregon, Eugene, OR 97403, USA}
%\author{Alexis Vives}
%\affiliation{Institute  for  Fundamental  Science, Department of Physics, University of Oregon, Eugene, OR 97403, USA}
%\author{Jon Hornung}
%\affiliation{Institute  for  Fundamental  Science, Department of Physics, University of Oregon, Eugene, OR 97403, USA}
%\affiliation{Thurston High School, Springfield, OR 9747, USA}

\begin{abstract}                 
We introduce a non-parametric model built from basis splines and use it to infer the astrophysical distribution of 
binary black hole (BBH) masses and spins using the 69 confident observations in GWTC-3, which marks this work as 
the most comprehensive data-driven investigation of the BBH population to date. We report the same features previously recovered 
with similarly flexible models focusing solely on the mass distribution, most notably the peaks in merger rates at $\sim10\msun$ and $\sim35\msun$. 
Our study provides the first non-parametric investigation into BBH spin distribution, while simultaneously modeling the mass distribution in the same way, 
showing signs that the spin tilt distributions may not peak at $\cos(\theta_i) = 1$ as previous models assumed. We find broad agreement with the 
accepted conclusions of the spin magnitude distribution: the majority of BBH spins are small ($a<0.5$), the distribution peaks at $a\sim0.2$, 
and there is mild support for a non-spinning subpopulation, which may become resolvable with larger catalogs. 
We perform a data-driven test to determine if the data prefers differing spin distributions for each binary component, and 
find no strong sign that the distributions differ. When modeling the $\chi_\mathrm{eff}$ distribution with our model we find similar 
conclusions as previously reported: the distribution is consistent with a non-isotropic distribution (that peaks at $\chi_\mathrm{eff}=\,$\result{$\CIPlusMinus{\macros[ChiEffective][chieff][PeakChiEff]}$}), 
showing considerable support ($f_{\chi_\mathrm{eff}<0}=\,$\result{$\CIPlusMinus{\macros[ChiEffective][chieff][FracBelow0]}$}) for negative effective spins or misaligned systems.
We conclude with a discussion of how non-parametric methods in gravitational-wave population inference are uniquely poised to complement to 
the parametric approach as we enter the data-rich era of gravitational-wave astronomy.
\end{abstract}

@@

\section{Acknowledgements}\label{sec:acknowledments}
We thank Will Farr, Tom Callister, and Maya Fishbach for useful discussions during the preparation of this manuscript, 
along with helpful comments on an early draft by Tom Callister and Jaxen Godfrey.
This research has made use of data, software and/or web tools obtained from the Gravitational Wave Open Science Center 
(\url{https://www.gw-openscience.org/}), a service of LIGO Laboratory, the LIGO Scientific Collaboration and the Virgo Collaboration. 
The authors are grateful for computational resources provided by the LIGO Laboratory and supported by National Science Foundation Grants PHY-0757058 and PHY-0823459.  
This work benefited from access to the University of Oregon high performance computer, Talapas. This material is based upon work supported 
in part by the National Science Foundation under Grant PHY-1807046 and work supported by NSF's LIGO Laboratory which is a major facility 
fully funded by the National Science Foundation.
\software{
\textsc{Astropy}~\citep{2018AJ....156..123A},
\textsc{NumPy}~\citep{harris2020array},
\textsc{Matplotlib}~\citep{Hunter:2007},
\textsc{Jax}~\citep{jax},
\textsc{NumPyro}~\citep{pyro,numpyro},
\textsc{GWPopulation}~\citep{Talbot_2019},
\textsc{Bilby}~\citep{Ashton_2019,bilby_gwtc1},
\textsc{Dynesty}~\citep{Speagle_2020}
}
\bibliography{bib}{}
\bibliographystyle{aasjournal}
